### 2018-11-07 articleSFDS.tex
ajout des nouveaux résultats

As defined Section \ref{sec:calibration}, when the code is time consuming, the solution is to mimic it with a Gaussian process (GP). For the GP emulator, we made the choice to consider the mean function $m_S(\bullet,\bullet)$ as a linear combination of linear functions. That means $\textbf{H}_S$ is a matrix of linear functions. The correlation function $r_S$ ($c_S=\sigma_S^2r_S$) chosen is defined by the following equation that corresponds to a Gaussian kernel:

@@

% \begin{equation*}
% \forall i \in \llbracket1,n\rrbracket \quad y_{exp_i}=F(\boldsymbol{x}_i,a_r)+\epsilon(\boldsymbol{x}_i)
% \end{equation*}
% where $\epsilon(\boldsymbol{x}_i)\sim\mathcal{N}(0,\sigma_{err}^2$ and $F(.,.)\sim\mathcal{GP}(m_1(.,.),c_{1_{\boldsymbol{\psi}_1}}(\{.,.\},\{.,.\})$.\newline

\begin{equation}
r_S\{(\boldsymbol{x},\boldsymbol{\theta}),(\boldsymbol{x}^*,\boldsymbol{\theta}^*)\}= \exp\Big\{-\frac{1}{2}\frac{||(\boldsymbol{x},\boldsymbol{\theta})-(\boldsymbol{x}^*,\boldsymbol{\theta}^*)||_2^2}{\boldsymbol{\psi}_S^2}\Big\}
\label{eq:GaussianKer}
\end{equation}

@@

The third model introduces another GP for the discrepancy. We choose a different covariance kernel which is a Matérn 5/2 (Equation (\ref{eq:Matern52Ker})).
Note that compared to Equation (\ref{eq:M3}), the discrepancy mean has been set to $0$ (\textit{i.e.} $m_{\boldsymbol{\delta}}(.)=0$). These choices are motivated by the fact that the purpose of calibration is to estimate the "best-fitting" 
vector parameter $\boldsymbol{\theta}$. We do not want any compensation into any additional bias. This decision is consistent with \citet{bachoc2014} where the same hypothesis has been made.

%The application of the third model is motivated by the assumption that there might occur a gap between the code and the phenomenon. Even after considering the measurement error, this gap is expected. This error term is called code error or discrepancy $\delta(\bullet)$. In a case where the code is not time consuming, the model which deals with this situation is $\mathcal{M}_3$ (Equation (\ref{eq:M3})). In the general framework $\delta(\bullet)\sim\mathcal{PG}(m_{\boldsymbol{\delta}}(\bullet),c_{\boldsymbol{\delta}}(\bullet,\bullet))$.\newline

@@

\begin{equation}
r_{\delta}(\boldsymbol{x},\boldsymbol{x}^*)= \Big(1+\frac{\sqrt{5}||\boldsymbol{x}-\boldsymbol{x}^*||_2}{\boldsymbol{\psi}_{\delta}}+\frac{5||\boldsymbol{x}-\boldsymbol{x}^*||_2^2}{3\boldsymbol{\psi}_{\delta}^2}\Big)\exp\Big\{-\frac{\sqrt{5}||\boldsymbol{x}-\boldsymbol{x}^*||_2}{\boldsymbol{\psi}_{\delta}}\Big\}.
\label{eq:Matern52Ker}
\end{equation}

@@

\begin{figure}[htbp!]
\begin{center}
  \begin{tabular}{ccccc}
&  $\mathcal{M}_1$ & $\mathcal{M}_2$ & $\mathcal{M}_3$ & $\mathcal{M}_4$ \\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    & \includegraphics[width=.2\textwidth]{figR/model1/densityEta.pdf} 
    &  \includegraphics[width=.2\textwidth]{figR/model2/densityEta.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model3/densityEta.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/densityEta.pdf}\\
	&\multicolumn{4}{c}{$\eta$}\\
	&&&&\\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    & \includegraphics[width=.2\textwidth]{figR/model1/densityMu.pdf} 
    &  \includegraphics[width=.2\textwidth]{figR/model2/densityMu.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model3/densityMu.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/densityMu.pdf}\\
	&\multicolumn{4}{c}{$\mu_t$}\\
	&&&&\\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    & \includegraphics[width=.2\textwidth]{figR/model1/densityAr.pdf} 
    &  \includegraphics[width=.2\textwidth]{figR/model2/densityAr.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model3/densityAr.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/densityAr.pdf}\\
	&\multicolumn{4}{c}{$a_r$}\\
	&&&&\\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    & \includegraphics[width=.2\textwidth]{figR/model1/densitySig.pdf} 
    &  \includegraphics[width=.2\textwidth]{figR/model2/densitySig.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model3/densitySig.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/densitySig.pdf}\\
	&\multicolumn{4}{c}{$\sigma_{err}^2$}\\
  \end{tabular}
\caption{\textit{Prior} (in blue) and posterior (in red) densities of $\eta$, $\mu_t$, $a_r$ and $\sigma_{err}^2$ for each model.
On the two first column the two first models (without and with surrogate) which have only these four parameters to estimate.
The two other columns represent the third and the fourth models which have two more parameters to estimate (see Figure \ref{fig:comparisionDensities2}).}
\label{fig:comparisionDensities1}
\end{center}
\end{figure}

\begin{figure}[htbp!]
\begin{center}
  \begin{tabular}{ccc}
& $\mathcal{M}_2'$ & $\mathcal{M}_4'$ \\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    &  \includegraphics[width=.2\textwidth]{figR/model2/after/densityEta.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/after/densityEta.pdf}\\
	&\multicolumn{2}{c}{$\eta$}\\
	&&\\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    &  \includegraphics[width=.2\textwidth]{figR/model2/after/densityMu.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/after/densityMu.pdf}\\
	&\multicolumn{2}{c}{$\mu_t$}\\
	&&\\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    &  \includegraphics[width=.2\textwidth]{figR/model2/after/densityAr.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/after/densityAr.pdf}\\
	&\multicolumn{2}{c}{$a_r$}\\
	&&\\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    &  \includegraphics[width=.2\textwidth]{figR/model2/after/densitySig.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/after/densitySig.pdf}\\
	&\multicolumn{2}{c}{$\sigma_{err}^2$}\\
  \end{tabular}
\caption{Calibration results for $\mathcal{M}_2'$ and $\mathcal{M}_4'$ which uses the surrogate based on the sequential design.}
\label{fig:calibrationSeq}
\end{center}
\end{figure}

@@

\begin{figure}[htbp!]
\begin{center}
  \begin{tabular}{cccc}
&  $\mathcal{M}_3$ & $\mathcal{M}_4$ & $\mathcal{M}_4'$ \\
    \rotatebox{90}{ \hspace{3em} \small density}
	&  \includegraphics[width=.2\textwidth]{figR/model3/densitySigD.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/densitySigD.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/after/densitySigD.pdf}\\
	&\multicolumn{3}{c}{$\sigma_{\boldsymbol{\delta}}^2$}\\
	&&&\\
    \rotatebox{90}{ \hspace{3em} \small density}
	&  \includegraphics[width=.2\textwidth]{figR/model3/densityPsiD.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/densityPsiD.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/after/densityPsiD.pdf}\\
	&\multicolumn{3}{c}{$\psi_{\boldsymbol{\delta}}$}\\
  \end{tabular}   
\caption{\textit{Prior} (in blue) and posterior (in red) densities of $\sigma_{\boldsymbol{\delta}}^2$ and $\psi_{\boldsymbol{\delta}}$ for $\mathcal{M}_3$, $\mathcal{M}_4$ and $\mathcal{M}_4'$.}
\label{fig:comparisionDensities2}
\end{center}
\end{figure}

>>>>

As defined Section \ref{sec:calibration}, when the code is time consuming, the solution is to mimic it with a Gaussian process (GP). For the GP emulator, we made the choice to consider the mean function $m_S(\bullet,\bullet)$ as a linear combination of linear functions. That means $\textbf{H}_S$ is a matrix of linear functions. The correlation function $r_S$ ($c_S=\sigma_S^2r_S$) chosen is defined by the following equation that corresponds to a Matérn $5/2$ kernel:

@@

% \begin{equation*}
% \forall i \in \llbracket1,n\rrbracket \quad y_{exp_i}=F(\boldsymbol{x}_i,a_r)+\epsilon(\boldsymbol{x}_i)
% \end{equation*}
% where $\epsilon(\boldsymbol{x}_i)\sim\mathcal{N}(0,\sigma_{err}^2$ and $F(.,.)\sim\mathcal{GP}(m_1(.,.),c_{1_{\boldsymbol{\psi}_1}}(\{.,.\},\{.,.\})$.\newline
\begin{equation}
r_{\delta}(\boldsymbol{x},\boldsymbol{x}^*)= \Big(1+\frac{\sqrt{5}||\boldsymbol{x}-\boldsymbol{x}^*||_2}{\boldsymbol{\psi}_{\delta}}+\frac{5||\boldsymbol{x}-\boldsymbol{x}^*||_2^2}{3\boldsymbol{\psi}_{\delta}^2}\Big)\exp\Big\{-\frac{\sqrt{5}||\boldsymbol{x}-\boldsymbol{x}^*||_2}{\boldsymbol{\psi}_{\delta}}\Big\}.
\label{eq:Matern52Ker}
\end{equation}

@@

The third model introduces another GP for the discrepancy. We choose a different covariance kernel which is Gaussian (Equation (\ref{eq:GaussianKer})).
Note that compared to Equation (\ref{eq:M3}), the discrepancy mean has been set to $0$ (\textit{i.e.} $m_{\boldsymbol{\delta}}(.)=0$). These choices are motivated by the fact that the purpose of calibration is to estimate the "best-fitting" 
vector parameter $\boldsymbol{\theta}$. We do not want any compensation into any additional bias. This decision is consistent with \citet{bachoc2014} where the same hypothesis has been made.

%The application of the third model is motivated by the assumption that there might occur a gap between the code and the phenomenon. Even after considering the measurement error, this gap is expected. This error term is called code error or discrepancy $\delta(\bullet)$. In a case where the code is not time consuming, the model which deals with this situation is $\mathcal{M}_3$ (Equation (\ref{eq:M3})). In the general framework $\delta(\bullet)\sim\mathcal{PG}(m_{\boldsymbol{\delta}}(\bullet),c_{\boldsymbol{\delta}}(\bullet,\bullet))$.\newline
\begin{equation}
r_S\{(\boldsymbol{x},\boldsymbol{\theta}),(\boldsymbol{x}^*,\boldsymbol{\theta}^*)\}= \exp\Big\{-\frac{1}{2}\frac{||(\boldsymbol{x},\boldsymbol{\theta})-(\boldsymbol{x}^*,\boldsymbol{\theta}^*)||_2^2}{\boldsymbol{\psi}_S^2}\Big\}
\label{eq:GaussianKer}
\end{equation}

@@

@@

\begin{figure}[htbp!]
\begin{center}
  \begin{tabular}{ccccc}
&  $\mathcal{M}_1$ & $\mathcal{M}_2$ & $\mathcal{M}_3$ & $\mathcal{M}_4$ \\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    & \includegraphics[width=.2\textwidth]{new/Model1/eta.pdf}
    &  \includegraphics[width=.2\textwidth]{new/Model2/eta.pdf}
	&  \includegraphics[width=.2\textwidth]{new/Model3/eta.pdf}
	&  \includegraphics[width=.2\textwidth]{new/Model4/eta.pdf}\\
	&\multicolumn{4}{c}{$\eta$}\\
	&&&&\\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    & \includegraphics[width=.2\textwidth]{new/Model1/mu.pdf}
    &  \includegraphics[width=.2\textwidth]{new/Model2/mu.pdf}
	&  \includegraphics[width=.2\textwidth]{new/Model3/mu.pdf}
	&  \includegraphics[width=.2\textwidth]{new/Model4/mu.pdf}\\
	&\multicolumn{4}{c}{$\mu_t$}\\
	&&&&\\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    & \includegraphics[width=.2\textwidth]{new/Model1/ar.pdf} 
    &  \includegraphics[width=.2\textwidth]{new/Model2/ar.pdf}
	&  \includegraphics[width=.2\textwidth]{new/Model3/ar.pdf}
	&  \includegraphics[width=.2\textwidth]{new/Model4/ar.pdf}\\
	&\multicolumn{4}{c}{$a_r$}\\
	&&&&\\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    & \includegraphics[width=.2\textwidth]{new/Model1/Serr.pdf} 
    &  \includegraphics[width=.2\textwidth]{new/Model2/Serr.pdf}
	&  \includegraphics[width=.2\textwidth]{new/Model3/Serr.pdf}
	&  \includegraphics[width=.2\textwidth]{new/Model4/Serr.pdf}\\
	&\multicolumn{4}{c}{$\sigma_{err}^2$}\\
  \end{tabular}
\caption{\textit{Prior} (in blue) and posterior (in red) densities of $\eta$, $\mu_t$, $a_r$ and $\sigma_{err}^2$ for each model.
On the two first column the two first models (without and with surrogate) which have only these four parameters to estimate.
The two other columns represent the third and the fourth models which have two more parameters to estimate (see Figure \ref{fig:comparisionDensities2}).}
\label{fig:comparisionDensities1}
\end{center}
\end{figure}

\begin{figure}[htbp!]
\begin{center}
  \begin{tabular}{ccc}
& $\mathcal{M}_2'$ & $\mathcal{M}_4'$ \\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    &  \includegraphics[width=.2\textwidth]{new/Model2sd/eta.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/after/densityEta.pdf}\\
	&\multicolumn{2}{c}{$\eta$}\\
	&&\\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    &  \includegraphics[width=.2\textwidth]{new/Model2sd/mu.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/after/densityMu.pdf}\\
	&\multicolumn{2}{c}{$\mu_t$}\\
	&&\\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    &  \includegraphics[width=.2\textwidth]{new/Model2sd/ar.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/after/densityAr.pdf}\\
	&\multicolumn{2}{c}{$a_r$}\\
	&&\\
    \rotatebox{90}{ \hspace{3em} \footnotesize density}
    &  \includegraphics[width=.2\textwidth]{new/Model2sd/Serr.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/after/densitySig.pdf}\\
	&\multicolumn{2}{c}{$\sigma_{err}^2$}\\
  \end{tabular}
\caption{Calibration results for $\mathcal{M}_2'$ and $\mathcal{M}_4'$ which uses the surrogate based on the sequential design.}
\label{fig:calibrationSeq}
\end{center}
\end{figure}

@@

\begin{figure}[htbp!]
\begin{center}
  \begin{tabular}{cccc}
&  $\mathcal{M}_3$ & $\mathcal{M}_4$ & $\mathcal{M}_4'$ \\
    \rotatebox{90}{ \hspace{3em} \small density}
	&  \includegraphics[width=.2\textwidth]{new/Model3/Serr.pdf}
	&  \includegraphics[width=.2\textwidth]{new/Model4/Sig.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/after/densitySigD.pdf}\\
	&\multicolumn{3}{c}{$\sigma_{\boldsymbol{\delta}}^2$}\\
	&&&\\
    \rotatebox{90}{ \hspace{3em} \small density}
	&  \includegraphics[width=.2\textwidth]{new/Model3/psi.pdf}
	&  \includegraphics[width=.2\textwidth]{new/Model4/psi.pdf}
	&  \includegraphics[width=.2\textwidth]{figR/model4/after/densityPsiD.pdf}\\
	&\multicolumn{3}{c}{$\psi_{\boldsymbol{\delta}}$}\\
  \end{tabular}   
\caption{\textit{Prior} (in blue) and posterior (in red) densities of $\sigma_{\boldsymbol{\delta}}^2$ and $\psi_{\boldsymbol{\delta}}$ for $\mathcal{M}_3$, $\mathcal{M}_4$ and $\mathcal{M}_4'$.}
\label{fig:comparisionDensities2}
\end{center}
\end{figure}
